### 2016-01-23 ijcar-abts.tex
IJCAR version: import body & appendices

\documentclass{llncs}

\usepackage[bw]{notation/modes}
\usepackage{notation/judgments}
\usepackage{notation/abts}
\usepackage{notation/common}

\usepackage{appendix}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bbold}
% \usepackage{amsthm}
\usepackage{catchfilebetweentags}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{ifdraft}
\usepackage{ifthen}
\usepackage{mathtools}
\usepackage{proof}
\usepackage{scalerel}
\usepackage{setspace}
\usepackage{stmaryrd}
\usepackage{supertabular}
\usepackage{tikz-cd}
\usepackage[length=.5]{tikzpfeile}
\usepackage{todonotes}
\usepackage{url}
\usepackage[h]{esvect}
\usepackage{ssabt}

\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\numberwithin{equation}{section}

\renewcommand\vec[1]{\protect\vv{#1}}
\renewcommand\MkAbs[3]{\{#1\}[#2].\, #3}

\begin{document}

\title{Syntax and Semantics of Abstract Binding Trees}
\date{}
\author{Jon Sterling and Darin Morrison}
\institute{}
\maketitle

\input{body}

\appendix
\appendixpage
\input{appendices}

\ifdraft{}{
  \newpage
  % \nocite{*}
  \bibliographystyle{abbrv}
  \bibliography{refs}
}

\end{document}